---
title: "ACS Diversity"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

This is a brief script to calculate the diversity of micro- and metro-SAs using ACS data, the R tidycensus library, and Simpson's Diversity Index via the R vegan library. Then it shows the top 10 MSAs by their SDI. Then we compute some SDIs of other geographic divisions.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidycensus)
library(vegan)
library(knitr)
```

Now get some values for population by ethnicity. The variable names represent the following:

* B03002_001 - Total
* B03002_003 - White Not Hispanic
* B03002_004 - Black Not Hispanic
* B03002_005 - American Indian and Alask Native Not Hispanic
* B03002_006 - Asian Not Hispanic
* B03002_007 - Native Hawaiian and Other Pacific Islander Not Hispanic
* B03002_008 - Some Other Race Not Hispanic
* B03002_009 - Two or More Races Not Hispanic
* B03002_012 - Hispanic or Latino

```{r message=FALSE}
race <- get_acs(geography = "metropolitan statistical area/micropolitan statistical area", variables = c("B03002_001","B03002_003","B03002_004","B03002_005","B03002_006","B03002_007","B03002_008","B03002_009","B03002_012"))
race <- select(race,-c(moe))
races <- spread(race,variable,estimate)
```

Next calculate the Simpson Diversity Index, or the Herfindahl–Hirschman index (HHI). Essentially, this index measures the probability that two specimens taken from the population will be of different type.

```{r message=FALSE}
races$sdi <- diversity(races[,4:11],"simpson")
races %>%
  top_n(10) %>%
  arrange(desc(sdi)) %>%
  select(NAME, B03002_001, sdi) %>%
  kable(digits=3, col.names = c("MSA","Total population","SDI"), format.args = list(big.mark = ","))
```

Lots of places with a fairly small population. What about the top 10 over some arbitrary total, like a million and one people?

```{r message=FALSE}
races %>%
  filter(B03002_001 >= 1000001) %>%
  top_n(10) %>%
  arrange(desc(sdi)) %>%
  select(NAME, B03002_001, sdi) %>%
  kable(digits=3, col.names = c("MSA","Total population","SDI"), format.args = list(big.mark = ","))
```

Let's take a look by county and do it all again.

```{r message=FALSE}
race <- get_acs(geography = "county", variables = c("B03002_001","B03002_003","B03002_004","B03002_005","B03002_006","B03002_007","B03002_008","B03002_009","B03002_012"))
race <- select(race,-c(moe))
races <- spread(race,variable,estimate)
races$sdi <- diversity(races[,4:11],"simpson")
races %>%
  filter(B03002_001 >= 200001) %>%
  top_n(10) %>%
  arrange(desc(sdi)) %>%
  select(NAME, B03002_001, sdi) %>%
  kable(digits=3, col.names = c("County","Total population","SDI"), format.args = list(big.mark = ","))
```

Or congressional districts with the highest diversity:

```{r message=FALSE}
race <- get_acs(geography = "congressional district", variables = c("B03002_001","B03002_003","B03002_004","B03002_005","B03002_006","B03002_007","B03002_008","B03002_009","B03002_012"))
race <- select(race,-c(moe))
races <- spread(race,variable,estimate)
races$sdi <- diversity(races[,4:11],"simpson")
races %>%
  #filter(B03002_001 >= 200001) %>%
  top_n(10) %>%
  arrange(desc(sdi)) %>%
  select(NAME, B03002_001, sdi) %>%
  kable(digits=3, col.names = c("Congressional district","Total population","SDI"), format.args = list(big.mark = ","))
```

and the districts with the lowest:

```{r message=FALSE}
races %>%
  top_n(-10) %>%
  arrange(sdi) %>%
  select(NAME, B03002_001, sdi) %>%
  kable(digits=3, col.names = c("Congressional district","Total population","SDI"), format.args = list(big.mark = ","))
```

How does diversity of a congressional district compare to its political leaning? Let's use data from the [Cook Political Report](https://cookpolitical.com/pvi-map-and-district-list). First we need to do some manipulation to get the Census's district IDs to match Cook's in order to properly join the tables.

```{r warning=FALSE, message = FALSE}
races$nums <- str_extract(races$NAME, " \\d\\d? ") %>% str_trim()
races$nums <- sprintf("%02d", as.numeric(races$nums))
races$nums[is.na(races$nums)] <- "AL"
races <- races %>%
  mutate(states = state.abb[match(str_extract(NAME, '\\b[^,]+$'),state.name)]) %>%
  mutate(Dist = str_c(states, nums, sep = "-"))
cook <- read.csv("data-5vPn3.csv", header=TRUE)
cds <- inner_join(cook, races)
cds %>%
  top_n(20, sdi) %>%
  arrange(desc(sdi)) %>%
  select(NAME, Incumbent, B03002_001, sdi, PVI) %>%
  kable(digits=3, col.names = c("Congressional district","Member","Total population","SDI","PVI"), format.args = list(big.mark = ","))
```

Most, but not all, diverse districts lean Democratic. What about the 20 least diverse districts?

```{r}
cds %>%
  top_n(-20, sdi) %>%
  arrange(sdi) %>%
  select(NAME, Incumbent, B03002_001, sdi, PVI) %>%
  kable(digits=3, col.names = c("Congressional district","Member","Total population","SDI","PVI"), format.args = list(big.mark = ","))
```
