---
title: "ACS Diversity"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

This is a brief script to calculate the diversity of micro- and metro-SAs using ACS data, the R tidycensus library, and Simpson's Diversity Index via the R vegan library. Then it shows the top 10 MSAs by their SDI.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidycensus)
library(vegan)
library(knitr)
```

Now get some values for population by ethnicity. The variable names represent the following:

* B03002_001 - Total
* B03002_003 - White Not Hispanic
* B03002_004 - Black Not Hispanic
* B03002_005 - American Indian and Alask Native Not Hispanic
* B03002_006 - Asian Not Hispanic
* B03002_007 - Native Hawaiian and Other Pacific Islander Not Hispanic
* B03002_008 - Some Other Race Not Hispanic
* B03002_009 - Two or More Races Not Hispanic
* B03002_012 - Hispanic or Latino

```{r message=FALSE}
race <- get_acs(geography = "metropolitan statistical area/micropolitan statistical area", variables = c("B03002_001","B03002_003","B03002_004","B03002_005","B03002_006","B03002_007","B03002_008","B03002_009","B03002_012"))
race <- select(race,-c(moe))
races <- spread(race,variable,estimate)
```

Next calculate the Simpson Diversity Index, or the Herfindahl–Hirschman index (HHI). Essentially, this index measures the probability that two specimens taken from the population will be of different type.

```{r message=FALSE}
races$sdi <- diversity(races[,4:11],"simpson")
races %>%
  top_n(10) %>%
  arrange(desc(sdi)) %>%
  select(NAME, B03002_001, sdi) %>%
  kable(digits=3, col.names = c("MSA","Total population","SDI"), format.args = list(big.mark = ","))
```

Lots of places with a fairly small population. What about the top 10 over some arbitrary total, like a million and one people?

```{r message=FALSE}
races %>%
  filter(B03002_001 >= 1000001) %>%
  top_n(10) %>%
  arrange(desc(sdi)) %>%
  select(NAME, B03002_001, sdi) %>%
  kable(digits=3, col.names = c("MSA","Total population","SDI"), format.args = list(big.mark = ","))
```

